### 2013-01-31 mcmc.tex
small tweaks and notes to self

\noindent
Dan~Foreman-Mackey\\
\affil{Center~for~Cosmology~and~Particle~Physics, Department~of~Physics,%
       New York University}
\\[1ex]
David~W.~Hogg\\
\affil{Center~for~Cosmology~and~Particle~Physics, Department~of~Physics,%
       New York University}\\
\affil{Max-Planck-Institut f\"ur Astronomie, Heidelberg}

@@

Say you are in this situation.
You have a huge blob of data $\data$ (think of this as a vector or
list or heterogeneous collection of observations\note{In general in
  this \documentname\ we will treat any variable as being capable of
  containing not just a value but some huge arbitrarily structured set
  of values in matrix, vector, or more complex form.}).
You also have a model sophisticated enough---a probabilistic,
generative model, if you will\note{For our definition of a
  \emph{model}, see [SOME OTHER NOTE] in this series.}---that, given a
setting of a huge blob of parameters $\pars$\note{Again, the parameter
  blob $\pars$ should also be thought of as an immense, arbitrarily
  structured set or list of parameters, possibly even with no fixed
  dimension!}, you can compute a pdf for data (or likelihood\note{We
  bash the terminology ``likelihood'' in [SOME OTHER NOTE] in this
  series.}) $p(\data\given\pars)$.
% ************* BJB
% <Wild Subjective Bayesian>
% The "generative model" is a probabilistic description of your *prior
% beliefs about what the data are going to be*. </Wild Subjective Bayesian>
And say you also can write down some kind of informative or
vague prior pdf $p(\pars)$ for the parameter blob.
If all these things are true, then---even if you can't compute
anything else---in principle a trivial-to-implement MCMC can give you
a fair sampling of the posterior pdf.
That is, you can run MCMC (for a very long time\note{Sometimes a very,
  very long time; we will return to that below.})  and you will be
left with a set of $K$ parameter-blob settings $\pars_k$ such that the
full set $\setofall{\pars_k}_{k=1}^K$ constitutes a fair sampling from
the posterior pdf $p(\pars\given\data$).
% ************* BJB
% Notions of "sampling" and "Monte Carlo" are interesting. You might like to
% refer the interested reader to the paper "Monte Carlo is Fundamentally
% Unsound" by Tony O'Hagan.

>>>>

\noindent
Dan~Foreman-Mackey\\
\affil{Center~for~Cosmology~and~Particle~Physics, Department~of~Physics,
       New York University}\\[1ex]
David~W.~Hogg\\
\affil{Center~for~Cosmology~and~Particle~Physics, Department~of~Physics,
       New York University}\\
\affil{Max-Planck-Institut f\"ur Astronomie, Heidelberg}

@@

Say you are in this situation.
You have a huge blob of data $\data$ (think of this as a vector or
list or heterogeneous collection of observations\note{In general in
  this \documentname\ we will treat any variable as being capable of
  containing not just a value but some huge arbitrarily structured set
  of values in matrix, vector, or more complex form.}).
You also have a model sophisticated enough---a probabilistic,
generative model, if you will\note{For our definition of a
  \emph{model}, see [SOME OTHER NOTE] in this series.}---that, given a
setting of a huge blob of parameters\note{Again, the parameter
  blob $\pars$ should also be thought of as an immense, arbitrarily
  structured set or list of parameters, possibly even with no fixed
  dimension!} $\pars$, you can compute a pdf for data (or likelihood\note{We
  bash the terminology ``likelihood'' in [SOME OTHER NOTE] in this
  series.}) $p(\data\given\pars)$.
% ************* BJB
% <Wild Subjective Bayesian>
% The "generative model" is a probabilistic description of your *prior
% beliefs about what the data are going to be*. </Wild Subjective Bayesian>
% ************* Hogg
% That comment should go in the ``What is a Model?'' Chapter.
And say you also can write down some kind of informative or
vague prior pdf $p(\pars)$ for the parameter blob.
If all these things are true, then---even if you can't compute
anything else---in principle a trivial-to-implement MCMC can give you
a fair sampling of the posterior pdf.
That is, you can run MCMC (for a very long time\note{Sometimes a very,
  very long time; we will return to that below.})  and you will be
left with a set of $K$ parameter-blob settings $\pars_k$ such that the
full set $\setofall{\pars_k}_{k=1}^K$ constitutes a fair sampling from
the posterior pdf $p(\pars\given\data$).
% ************* BJB
% Notions of "sampling" and "Monte Carlo" are interesting. You might like to
% refer the interested reader to the paper "Monte Carlo is Fundamentally
% Unsound" by Tony O'Hagan.
